% === [ Conclusion ] ===========================================================

\section{Conclusion}

This section concludes the project report and includes subjective reflections from the author. For the remainder of this section I will switch to a first person narrative.

% --- [ Project Summary ] ------------------------------------------------------

\subsection{Project Summary}

% TODO: Summarise the key findings of your report. No new information should be included.

foo

% --- [ Personal Development ] -------------------------------------------------

\subsection{Personal Development}

This is the largest project I have undertaken in my life and I feel satisfied with the outcome and proud of what I have been able to accomplished. It has re-enforced my belief that any problem is solvable when broken into smaller subproblems and instilled me with a feeling that anything is possible. The project has allowed me to mature as a software developer and I now feel more confident in utilizing best practies such as TDD, CI and semantic versioning. I have also matured as a developer and gained experience with implementing a semi-large project and structuring it into several smaller self-contained projects.

% --- [ Final Thoughts ] -------------------------------------------------------

\subsection{Final Thoughts}

My happiest moment during the project was when the larger components started working and could be connected to form a complete system. It feels great having started out with a vague idea of how the decompiler could work, gradually gaining new insights and refining its design after researching and building on the knowledge of others, developing and iteratively reimplementing the various components until they feel just right, finally arriving at a working prototype and seeing the full system in action! If there is one key idea I want to leave you with it is that the composition of independent components, each with a single purpose and well-defined input and output, is a powerful concept for solving complex problems.
