% --- [ Personal Development ] -------------------------------------------------

\subsection{Personal Development}

This is the largest project I have undertaken in my life and I feel satisfied with the outcome and proud of what I have been able to accomplished. It has re-enforced my belief that any problem is solvable when broken into smaller subproblems and instilled me with a feeling that anything is possible. The project has allowed me to mature as a software engineer and I now feel more confident in utilizing best practices such as test-driven development and CI. I have also matured as a developer and gained experience with implementing a semi-large project and structuring it into several smaller self-contained projects.

If I were to redo the project today, I wish someone would have told me about the critical importance of effective time management, and then told me again to let it sink in. In the beginning of the project I used coarse time management and assigned bi-weekly deadlines to larger tasks. While some tasks were completed ahead of schedule, many were postponed. Later on it became apparent that accurate time estimates could only be given when larger tasks were broken down into sufficiently small sub-tasks. Towards the end of the project I used more granular time management, prioritised tasks, and split larger tasks into sub-tasks that could be completed in a couple of days. For software development tasks the former technique (coarse time management) worked well and for report writing tasks the latter technique (granular time management) was most effective.
