% --- [ Disposition ] ----------------------------------------------------------

\subsection{Disposition}

This report details every stage of the project from conceptualisation to successful completion. It follows a logical structure and outlines the major stages in chronological order. A brief summary of each section is presented in the list below.

\begin{itemize}
	\item Section \ref{sec:introduction} - \textbf{Introduction} \\ \textit{Introduces the concept of decompilation and its applications, outlines the project aim and objectives, and summarises its deliverables.}
	\item Section \ref{sec:literature_review} - \textbf{Literature Review} \\ \textit{Details the problem domain, reviews traditional decompilation techniques, and evaluates potential intermediate representations for the decompilation pipeline of the project.}
	\item Section \ref{sec:related_work} - \textbf{Related Work} \\ \textit{Evaluates projects for translating native code to LLVM IR, and reviews the design of modern decompilers.}
	\item Section \ref{sec:methodology} - \textbf{Methodology} \\ \textit{Surveys methodologies and best practices for software construction, and relates them to the specific problem domain.}
	\item Section \ref{sec:requirements} - \textbf{Requirements} \\ \textit{Specifies and prioritises the requirements of the project artefacts.}
	\item Section \ref{sec:design} - \textbf{Design} \\ \textit{Discusses the system architecture and the design of each component, motivates the choice of core algorithms and data structures, and highlights strengths and limitations of the design.}
	\item Section \ref{sec:implementation} - \textbf{Implementation} \\ \textit{Discusses language considerations, describes the implementation process, and showcases how set-backs were dealt with.}
	\item Section \ref{sec:verification} - \textbf{Verification} \\ \textit{Describes the approaches taken to validate the correctness, performance and security of the artefacts.}
\clearpage % <layout>
	\item Section \ref{sec:evaluation} - \textbf{Evaluation} \\ \textit{Assesses the outcome of the project and evaluates the artefacts against the requirements.}
	\item Section \ref{sec:conclusion} - \textbf{Conclusion} \\ \textit{Summarises the project outcomes, presents ideas for future work, reflects on personal development, and concludes with an attribution to the key idea of this project.}
\end{itemize}
